Stability and Performance of Coalitions of Prosumers Through Diversification in the Smart Grid

### 2016-02-18 Article.tex
update
corrections Monique

In this paper, we consider agents in the distribution network that own both renewable distributed energy resources (DER) and loads. We will refer to these agents as prosumers \cite{Rathnayaka2012} in the future. If a prosumer produces more than he consumes, he might be willing to sell power. We thus assume the existence of an energy market where producers post a production capacity for an upcoming period, and buyers decide what bids to purchase. When the production is stochastic, as it can be with DER, this approach gets more complicated and risk should be quantified and taken into account. We thus assume the existence of a market operator who constrains the market entrance. As single agents might not be productive or reliable enough, aggregators managing large portfolios of prosumers are more likely to enter such markets. While there exists several models for fixing prices and organizing bids in such stochastic settings, the present work does not focus on transactions on the market. Instead, we study how aggregators can manage their portfolios as to maximize the production that they offer, given that such production is stochastic by nature. A key point of this work is then to merge the interests of the market operator and aggregators in a single utility function. While the aggregators intend to maximize their benefits with better contracts, the operator concerns are related to the quality and reliability of the market supply (more information about the scenario can be found in \cite{Europe} page 67-68). If a buyer purchases a contract, the corresponding aggregator has to inject the contracted amount of power at any time of the contract, exposing himself to financial penalties if he deviates. Since prosumers use renewables, we are concerned with the probability that the production falls below the contract value. In this case, storage or demand side managment might be solutions to avoid penalties, but this probability should still be maintained low.

In this paper, we decided to use a coalitional framework in order to model aggregators. It is known that diversification of the assets is a way of minimizing risks when building a portfolio. We thus expect more stable and predictable energy productions for aggreagtors than for single agents. Nevertheless, all coalitions are not equally stable or productive. Attention should thus be paid to the aggregation step. Recall that prosumers have both consumption and production components, each depending on location and time, meaning that there are complex underlying correlations between the agents. This is a central topic of this paper : given N prosumers, what coalitions should be formed so that the compromise between expected production and variability is optimized ?

@@

Because agents are susceptible to fail for diverse reasons, the propensity of the system to undertake these failures is critical \cite{Pahwa}. We therefore investigate in this paper the resilience of the coalitions when prosumers fail. Despite the fact that losing agents is usually detrimental to the coalitions, we will see that the coalitions formed with our algorithm tend to be less impacted by random failures.

@@

In this paper we studied how aggregations of prosumers could be authorized to sell their surplus of production on the energy market. By relying on the past values of the agents, we constrained the market entry to both sufficiently productive and stable coalitions. The power that a coalition is able to propose on the market is therefore related to production and stability. The correlations between the prosumers impact directly the volatility of the coalitions, which led us to seek uncorrelated aggregations of agents. We used a graph representation of the correlation relationships between the agents as a reduced landscape for the coalition formation. A greedy algorithm that starts with cliques of the "de-correlation" graph of the agents and makes local improvements offers a good compromise between speed and quality of the results. We compared these results with random samplings, and an opposite strategy that clusters correlated agents together. We showed that the coalitions resulting from our algorithm are able to provide more power to the grid with a lower volatility. Because of their better production over volatility ratios, these coalitions will tend to use less storage and waste less energy than more unstable coalitions. Because in real situations, agents are prone to failure, resilience is also an important criterion for the quality of the aggregations. We therefore studied how the coalitions remain on the market when their agents fail randomly. We showed that, in this situation, the coalitions resulting from our algorithm better withstand losses of agents. A possible direction for future work could be to use, in addition to real weather data, data from energy markets and aggregators as to propose a prosumer aggregation model closer to real conditions.

>>>>

In this paper, we consider agents in the distribution network that own both renewable distributed energy resources (DER) and loads. We will refer to these agents as prosumers \cite{Rathnayaka2012} in the future. If a prosumer produces more than he consumes, he might be willing to sell power. We thus assume the existence of an energy market where producers post a production capacity for an upcoming period, and buyers decide what bids to purchase. When the production is stochastic, as it can be with DER, this approach gets more complicated and risk should be quantified and taken into account. We thus assume the existence of a market operator who constrains the market entrance. As single agents might not be productive or reliable enough, aggregators managing large portfolios of prosumers are more likely to enter such markets. While there exist several models for fixing prices and organizing bids in such stochastic settings, the present work does not focus on transactions on the market. Instead, we study how aggregators can manage their portfolios as to maximize the production that they offer, given that such production is stochastic by nature. A key point of this work is then to merge the interests of the market operator and aggregators in a single utility function. While the aggregators intend to maximize their benefits with better contracts, the operator concerns are related to the quality and reliability of the market supply (more information about the scenario can be found in \cite{Europe} page 67-68). If a buyer purchases a contract, the corresponding aggregator has to inject the contracted amount of power at any time of the contract, exposing himself to financial penalties if he deviates. Since prosumers use renewables, we are concerned with the probability that the production falls below the contract value. In this case, storage or demand side managment might be solutions to avoid penalties, but this probability should still be maintained low.

We use a coalitional framework in order to model aggregators. It is known that diversification of the assets is a way of minimizing risks when building a portfolio. We thus expect more stable and predictable energy productions for aggregators than for single agents. Nevertheless, all coalitions are not equally stable or productive. Attention should thus be paid to the aggregation step. Recall that prosumers have both consumption and production components, each depending on location and time, meaning that there are complex underlying correlations between the agents. This is a central topic of this paper : given N prosumers, what coalitions should be formed so that the compromise between expected production and variability is optimized ?

@@

Because agents are susceptible to fail for diverse reasons, the propensity of the system to undertake these failures is critical \cite{Pahwa}. We also investigate in this paper the resilience of the coalitions when prosumers fail. Despite the fact that losing agents is usually detrimental to the coalitions, we will see that the coalitions formed with our algorithm tend to be less impacted by random failures.

@@

In this paper we studied how aggregations of prosumers could be authorized to sell their surplus of production on the energy market. By relying on the past values of the agents, we constrained the market entry to both sufficiently productive and stable coalitions. The power that a coalition is able to propose on the market is therefore related to production and stability. The correlations between the prosumers impact directly the volatility of the coalitions, which led us to seek uncorrelated aggregations of agents. We used a graph representation of the correlation relationships between the agents as a reduced landscape for the coalition formation. A greedy algorithm that starts with cliques of the "de-correlation" graph of the agents and makes local improvements offers a good compromise between speed and quality of the results. We compared these results with random samplings, and an opposite strategy that clusters correlated agents together. We showed that the coalitions resulting from our algorithm are able to provide more power to the grid with a lower volatility. Because of their better production over volatility ratios, these coalitions will tend to use less storage and waste less energy than more unstable coalitions. Because in real situations, agents are prone to failure, resilience is also an important criterion for the quality of the aggregations. We therefore studied how the coalitions remain on the market when their agents fail randomly. We showed that, in this situation, the coalitions resulting from our algorithm better withstand losses of agents. A possible direction for future work could be to use, in addition to real weather data, data from energy markets and aggregators' portfolios as to propose a prosumer aggregation model closer to real conditions.
